
## Load library

```{r}
library(tidyverse)

```

## Load dataset
```{r}
clinical_data <- read_tsv("msk_chord_2024_clinical_data (2).tsv")
cat("Total Rows:", nrow(clinical_data), "\n")
cat("Total Columns:", ncol(clinical_data), "\n")

```

## DATA CLEANING
```{r}
clean_df <- clinical_data %>%
  select(
    Age = `Current Age`, Sex, Race, Ethnicity,
    Cancer_Type = `Cancer Type`, Stage = `Stage (Highest Recorded)`,
    Tumor_Site = `Primary Tumor Site`, Sample_Type = `Sample Type`,
    Smoking = `Smoking History (NLP)`, Prior_Tx = `Prior Treatment to MSK (NLP)`,
    TMB = `TMB (nonsynonymous)`, MSI = `MSI Score`,
    FGA = `Fraction Genome Altered`, Purity = `Tumor Purity`,
    OS_Months = `Overall Survival (Months)`
  ) %>%
  mutate(
    across(c(Sex, Race, Ethnicity, Stage, Tumor_Site, Sample_Type, Smoking, Prior_Tx), 
           ~replace_na(as.character(.), "Not_Reported")),
    across(c(Sex, Race, Ethnicity, Stage, Tumor_Site, Sample_Type, Smoking, Prior_Tx), 
           ~ifelse(. %in% c("Unknown", "Pt Refused To Answer", ""), "Not_Reported", .)),
    
    Age = replace_na(as.numeric(Age), median(as.numeric(Age), na.rm=TRUE)),
    TMB = replace_na(as.numeric(TMB), median(as.numeric(TMB), na.rm=TRUE)),
    MSI = replace_na(ifelse(as.numeric(MSI) < 0, NA, as.numeric(MSI)), median(as.numeric(MSI), na.rm=TRUE)),
    FGA = replace_na(as.numeric(FGA), median(as.numeric(FGA), na.rm=TRUE)),
    Purity = replace_na(as.numeric(Purity), median(as.numeric(Purity), na.rm=TRUE))
  )

```

## Create Proxy Variable

```{r}
med_val <- median(clean_df$OS_Months, na.rm=TRUE)
clean_df$Response <- factor(ifelse(clean_df$OS_Months >= med_val, "Good", "Poor"))
```

## Numeric Variable Plots vs Response
```{r}
numeric_vars <- c("Age","TMB","MSI","FGA","Purity")
```

## Age vs Response
```{r}
ggplot(clean_df, aes(x = Response, y = Age, fill = Response)) +
  geom_boxplot(alpha = 0.7, outlier.color = "red") +
  scale_fill_manual(values = c("Good" = "#27AE60", "Poor" = "#C0392B")) +
  labs(title = "Age Distribution by Response") +
  theme_minimal()
```

```{r}
ggplot(clean_df, aes(x = Age, fill = Response)) +
  geom_density(alpha = 0.4) +
  scale_fill_manual(values = c("Good" = "#27AE60", "Poor" = "#C0392B")) +
  labs(title = "Density of Age by Response") +
  theme_minimal()
```
## TMB vs Response
```{r}
ggplot(clean_df, aes(x = Response, y = TMB, fill = Response)) +
  geom_boxplot(alpha = 0.7, outlier.color = "red") +
  scale_fill_manual(values = c("Good" = "#27AE60", "Poor" = "#C0392B")) +
  labs(title = "TMB Distribution by Response") +
  theme_minimal()
```
```{r}
ggplot(clean_df, aes(x = TMB, fill = Response)) +
  geom_density(alpha = 0.4) +
  scale_fill_manual(values = c("Good" = "#27AE60", "Poor" = "#C0392B")) +
  labs(title = "Density of TMB by Response") +
  theme_minimal()
```
## MSI vs Response
```{r}
ggplot(clean_df, aes(x = Response, y = MSI, fill = Response)) +
  geom_boxplot(alpha = 0.7, outlier.color = "red") +
  scale_fill_manual(values = c("Good" = "#27AE60", "Poor" = "#C0392B")) +
  labs(title = "MSI Distribution by Response") +
  theme_minimal()
```
```{r}
ggplot(clean_df, aes(x = MSI, fill = Response)) +
  geom_density(alpha = 0.4) +
  scale_fill_manual(values = c("Good" = "#27AE60", "Poor" = "#C0392B")) +
  labs(title = "Density of MSI by Response") +
  theme_minimal()
```
## FGA vs Response
```{r}
ggplot(clean_df, aes(x = Response, y = FGA, fill = Response)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = c("Good" = "#2980B9", "Poor" = "#F39C12")) +
  labs(title = "FGA Distribution by Response") +
  theme_minimal()
```

## Purity vs Response
```{r}
ggplot(clean_df, aes(x = Response, y = Purity, fill = Response)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = c("Good" = "#27AE60", "Poor" = "#C0392B")) +
  labs(title = "Purity Distribution by Response") +
  theme_minimal()
```
## Categorical Variable Plots vs Response
```{r}
categorical_vars <- c("Sex","Stage","Cancer_Type","Tumor_Site","Smoking","Prior_Tx")
```
## Sex vs Response
```{r}
prop_sex <- prop.table(table(clean_df$Sex, clean_df$Response), margin=1)
print(round(prop_sex, 2))

ggplot(clean_df, aes(x = Sex, fill = Response)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c("Good" = "#3498DB", "Poor" = "#E91E63")) +
  labs(title = "Proportion of Response by Sex", y = "Proportion") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
## ## Stage vs Response
```{r}
prop_stage <- prop.table(table(clean_df$Stage, clean_df$Response), margin=1)
print(round(prop_stage, 2))

ggplot(clean_df, aes(x = Stage, fill = Response)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c("Good" = "#27AE60", "Poor" = "#C0392B")) +
  labs(title = "Proportion of Response by Stage", y = "Proportion") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
## Cancer_Type vs Response
```{r}
prop_cancer <- prop.table(table(clean_df$Cancer_Type, clean_df$Response), margin=1)
print(round(prop_cancer, 2))

ggplot(clean_df, aes(x = Cancer_Type, fill = Response)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c("Good" = "#27AE60", "Poor" = "#C0392B")) +
  labs(title = "Proportion of Response by Cancer Type", y = "Proportion") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
## Tumor_Site vs Response
```{r}
prop_tumor <- prop.table(table(clean_df$Tumor_Site, clean_df$Response), margin=1)
print(round(prop_tumor, 2))

ggplot(clean_df, aes(x = Tumor_Site, fill = Response)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c("Good" = "#27AE60", "Poor" = "#C0392B")) +
  labs(title = "Proportion of Response by Tumor Site", y = "Proportion") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
## Smoking vs Response
```{r}
prop_smoke <- prop.table(table(clean_df$Smoking, clean_df$Response), margin=1)
print(round(prop_smoke, 2))

ggplot(clean_df, aes(x = Smoking, fill = Response)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c("Good" = "#27AE60", "Poor" = "#C0392B")) +
  labs(title = "Proportion of Response by Smoking History", y = "Proportion") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
## Prior_Tx vs Response
```{r}
prop_tx <- prop.table(table(clean_df$Prior_Tx, clean_df$Response), margin=1)
print(round(prop_tx, 2))

ggplot(clean_df, aes(x = Prior_Tx, fill = Response)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c("Good" = "#27AE60", "Poor" = "#C0392B")) +
  labs(title = "Proportion of Response by Prior Treatment", y = "Proportion") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
## Stage vs FGA
```{r}
ggplot(clean_df, aes(x = Stage, y = FGA, fill = Stage)) +
  geom_boxplot(alpha = 0.7, outlier.color = "red") +
  labs(title = "Genomic Instability (FGA) Across Tumor Stages",
       subtitle = "Does the genome become more unstable as cancer progresses?",
       x = "Cancer Stage", y = "Fraction Genome Altered (FGA)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Prior Treatment vs FGA


```{r}
ggplot(clean_df, aes(x = Prior_Tx, y = FGA, fill = Response)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = c("Good" = "#27AE60", "Poor" = "#C0392B")) +
  labs(title = "Impact of Prior Treatment on Genomic Stability and Response",
       x = "Prior Treatment to MSK", y = "Fraction Genome Altered (FGA)") +
  theme_minimal()
```


## Summary Statistics Table
```{r}
summary_table <- clean_df %>%
  group_by(Response) %>%
  summarise(
    Count = n(),
    Mean_Age = mean(Age, na.rm=TRUE),
    Median_TMB = median(TMB, na.rm=TRUE),
    Median_MSI = median(MSI, na.rm=TRUE),
    Median_FGA = median(FGA, na.rm=TRUE),
    Median_Purity = median(Purity, na.rm=TRUE)
  )

print(summary_table)
```

## Categorical variable proportion table
```{r}
## Proportion table for categorical variables
categorical_vars <- c("Sex","Stage","Cancer_Type","Tumor_Site","Smoking","Prior_Tx")
for (var in categorical_vars) {
  cat("\n\n---", var, "---\n")
  print(prop.table(table(clean_df[[var]], clean_df$Response), margin=2))
}

```

