## Load library

```{r}
library(readr)
library(dplyr)
library(ggplot2)

```

## Load dataset
```{r}
clinical_data <- read_tsv("msk_chord_2024_clinical_data (2).tsv")


```

## Data Cleaning And Pre-Processing

## Detect Missing Value
```{r}
colSums(is.na(clinical_data))
```


## Remove Rows with Critical Missing Values

```{r}
clinical_data_clean <- clinical_data %>%
  filter(
    !is.na(`Current Age`),
    !is.na(Sex),
    !is.na(`Cancer Type`),
    !is.na(`Stage (Highest Recorded)`),
    !is.na(`Overall Survival Status`)
  )
```


## Clean NLP-Derived fields

```{r}
clinical_data_clean <- clinical_data_clean %>%
  mutate(
    `Cancer Type` = trimws(tolower(`Cancer Type`)),
    `Stage (Highest Recorded)`= trimws(toupper(`Stage (Highest Recorded)`)),
    `Clinical Summary`=trimws(`Clinical Summary`),
    `Smoking History (NLP)`= trimws(`Smoking History (NLP)`),
    Sex = factor(Sex)
  )
```


## Create Treatment Response

```{r}
clinical_data_clean <- clinical_data_clean %>%
  mutate(
    response_score =
      ifelse(grepl("LIVING",`Overall Survival Status`, ignore.case = TRUE), 1, 0) +
      ifelse(`Clinical Summary`%in% c("Localized", "In Situ"), 1, 0) +
      ifelse(is.na(`Metastatic Site`), 1, 0) +
      ifelse(!is.na(`Fraction Genome Altered`) & `Fraction Genome Altered` < 0.2, 1, 0) +
      ifelse(grepl("Non-Smoker|Never",`Smoking History (NLP)`,ignore.case = TRUE), 1, 0),
    treatment_response = ifelse(response_score >= 3, "Good Response", "Poor Response"),
    treatment_response_bin = ifelse(treatment_response == "Good Response", 1, 0)
  )


```


## Subsetting

```{r}
subset(clinical_data_clean, Sex == "Female")
```

## EDA
```{r}
clinical_data_clean %>%
  group_by(`Stage (Highest Recorded)`, treatment_response) %>%
  summarise(count=n(),.groups="drop")
  
```


## Data Visualization

## Age vs Treatment Response
```{r}
ggplot(clinical_data_clean,aes(x = `Current Age`, fill = treatment_response)
       )+
  geom_histogram(position = "identity", alpha = 0.6, bins = 20)+
  scale_fill_manual(values = c("Good Response" = "blue", "Poor Response" = "gray")) +
  labs(title = "Age vs Treatment Response")
```
## Gender vs Treatment Response
```{r}
ggplot(
  clinical_data_clean,
       aes(x = Sex, fill = treatment_response)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c("Good Response" = "blue", "Poor Response" = "gray")) +
  labs(title = "Treatment Response by Gender",
       y = "Proportion")

```

## Tumor Stage vs Treatment Response
```{r}
ggplot(
  clinical_data_clean,
  aes(x = `Stage (Highest Recorded)`, fill = treatment_response)
)+
  geom_bar(position = "fill")+
  scale_fill_manual(values = c("Good Response" = "blue", "Poor Response" = "gray")) +
  labs(title = "Treatment Response by Tumor Stage")
```

##Create binary numeric response
```{r}
clinical_data_clean <- clinical_data_clean %>%
  mutate(
    treatment_response_bin = ifelse(treatment_response == "Good Response", 1, 0)
  )

```



## Train-Test Split
```{r}
set.seed(123)
index <- sample(seq_len(nrow(clinical_data_clean)),
                size = 0.7*nrow(clinical_data_clean))

train <- clinical_data_clean[index,]
test <- clinical_data_clean[-index,]
```
`

## Logistic Regression Model
```{r}
model <- glm(
  treatment_response_bin ~ `Current Age` + Sex + `Stage (Highest Recorded)`+ `Cancer Type`,
  data = train,
  family = binomial
)

summary(model)

```


##Pipe Operator(%>%)
```{r}
clinical_data_clean %>%
  filter(`Current Age` > 40) %>%
  select(`Current Age`, Sex, `Cancer Type`, `Stage (Highest Recorded)`, treatment_response) %>%
  mutate(age_group = ifelse(`Current Age` < 60, "Middle", "Older")) %>%
  arrange(`Stage (Highest Recorded)`)
```


## Summary(model)
```{r}
predicted_probs <- predict(model, newdata = test, type = "response")
predicted_class <- ifelse(predicted_probs > 0.5, 1, 0)
accuracy <- mean(predicted_class == test$treatment_response_bin)

cat("Final Model Accuracy:", round(accuracy * 100, 2), "%\n")
```



